### 2021-11-19 manuscript/fig_from_alignment_to_variation_graph.tex
other details; image placeholder

\vspace{-1em}
\begin{figure}[ht!]
%  \centering
    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={left,top},capbesidewidth=0.45\linewidth}}]{figure}[\FBwidth]
    {\caption{\FIXME{put high-resolution image} Building variation graphs from alignment graphs. Top: an alignment graph between a set of sequences,
        showing alignment edges in grey and precedence (path) edges in colors associated with each path. Bottom: the variation graph
        resulting from the application of the transitive closure process in \textit{seqwish}, where aligned segments have been
        compressed into a single node.
    }\label{fig:from_alignment_to_variation_graph}}
    %{\includegraphics[width=\linewidth]{fig/OUP_First_SBk_Bot_8401}} %\includegraphics[width=5cm]{name}}
    %\caption{Performance evaluation of \textit{odgi build} when translating a 90-haplotype graph of human chromosome 6 into ODGI's native format.}

\end{figure}
\vspace{-1em}