### 2022-01-05 manuscript/fig_induction.tex
maybe better

\documentclass{article}
\usepackage{tikz}
\usepackage{comment}
\usetikzlibrary{positioning}

\pgfdeclarelayer{bg}    % declare background layer
\pgfsetlayers{bg,main}  % set the order of the layers (main is the standard layer)

\begin{document}
\begin{tikzpicture}[auto,
    node distance= 10mm and 20mm,
    roundnode/.style={circle, draw=black!60, fill=gray!10, very thick, minimum size=7mm},
    squarednode/.style={rectangle, draw=black!60, fill=gray!10, very thick, minimum size=5mm},
    keynode/.style={rectangle, draw=black!60, fill=gray!10, thick, minimum size=1mm}
]
%Nodes
\node[squarednode] (seqA1) {G};
\node[squarednode] (seqA2) [right of=seqA1] {A};
\node[squarednode] (seqA3) [right of=seqA2] {T};
\node[squarednode] (seqA4) [right of=seqA3] {T};
\node[squarednode] (seqA5) [right of=seqA4] {A};
\node[squarednode] (seqA6) [right of=seqA5] {C};
\node[squarednode] (seqA7) [right of=seqA6] {A};

%\node[roundnode]        (uppercircle)       [above=of seqA1] {1};
%\node[squarednode]      (rightsquare)       [below=of seqA1] {3};
%\node[roundnode]        (lowercircle)       [below=of seqA1] {4};

\node[squarednode] (seqB1) [below of=seqA1] {G};
\node[squarednode] (seqB2) [right of=seqB1] {C};
\node[squarednode] (seqB3) [right of=seqB2] {T};
\node[squarednode] (seqB4) [right of=seqB3] {T};
\node[squarednode] (seqB5) [right of=seqB4] {A};
\node[squarednode] (seqB6) [right of=seqB5] {C};
\node[squarednode] (seqB7) [right of=seqB6] {G};

\node[squarednode] (seqC1) [below of=seqB1] {G};
\node[squarednode] (seqC2) [right of=seqC1] {A};
\node[squarednode] (seqC3) [right of=seqC2] {T};
\node[squarednode] (seqC4) [right of=seqC3] {C};
\node[squarednode] (seqC5) [right of=seqC4] {A};

%Lines
\draw[line width=1.0] (seqA1) -- (seqA2);
\draw[line width=1.0] (seqA2) -- (seqA3);
\draw[line width=1.0] (seqA3) -- (seqA4);
\draw[line width=1.0] (seqA4) -- (seqA5);
\draw[line width=1.0] (seqA5) -- (seqA6);
\draw[line width=1.0] (seqA6) -- (seqA7);

\draw[line width=1.0] (seqB1) -- (seqB2);
\draw[line width=1.0] (seqB2) -- (seqB3);
\draw[line width=1.0] (seqB3) -- (seqB4);
\draw[line width=1.0] (seqB4) -- (seqB5);
\draw[line width=1.0] (seqB5) -- (seqB6);
\draw[line width=1.0] (seqB6) -- (seqB7);

\draw[line width=1.0] (seqC1) -- (seqC2);
\draw[line width=1.0] (seqC2) -- (seqC3);
\draw[line width=1.0] (seqC3) -- (seqC4);
\draw[line width=1.0] (seqC4) -- (seqC5);

\begin{scope}[node distance=1.3cm]
\node[roundnode] (g2) [below right = 3cm and 1.5cm of seqA1] {A};
\node[roundnode] (g1) [below left of=g2] {G};
\node[roundnode] (g3) [below right of=g1] {C};
\node[roundnode] (g4) [above right of=g3] {T};
\node[roundnode] (g5) [right of=g4] {T};
\node[roundnode] (g6) [right of=g5] {A};
\node[roundnode] (g7) [right of=g6] {C};
\node[roundnode] (g8) [above right of=g7] {A};
\node[roundnode] (g9) [below right of=g7] {G};
\end{scope}

\draw[line width=1.0] (g1) to (g2);
\draw[line width=1.0] (g1) to (g3);
\draw[line width=1.0] (g2) to (g4);
\draw[line width=1.0] (g3) to (g4);
\draw[line width=1.0] (g4) to (g5);
\draw[line width=1.0] (g4) to [looseness=1] (g7);
\draw[line width=1.0] (g5) to (g6);
\draw[line width=1.0] (g6) to (g7);
\draw[line width=1.0] (g7) to (g8);
\draw[line width=1.0] (g7) to (g9);

\begin{pgfonlayer}{bg}    % select the background layer
\draw[line width=0.5,color=red!100,looseness=1] (seqA1.south) to (g1);
\draw[line width=0.5,color=red!100,looseness=1] (seqB1.south) to (g1);
\draw[line width=0.5,color=red!100,looseness=1] (seqC1.south) to (g1);

\draw[line width=0.5,color={rgb:red,1;orange,1},looseness=1.0] (seqA2.south) to (g2);
\draw[line width=0.5,color={rgb:red,1;orange,1},looseness=1.0] (seqC2.south) to (g2);

\draw[line width=0.5,color=orange!100,looseness=1.0] (seqA3.south) to (g4);
\draw[line width=0.5,color=orange!100,looseness=1.0] (seqB3.south) to (g4);
\draw[line width=0.5,color=orange!100,looseness=1.0] (seqC3.south) to (g4);

\draw[line width=0.5,color=green!100,looseness=1.0] (seqA4.south) to (g5);
\draw[line width=0.5,color=green!100,looseness=1.0] (seqB4.south) to (g5);
%\draw[line width=0.5,color=blue!100,looseness=1.0] (seqC3.south) to (g4);

\draw[line width=0.5,color={rgb:blue,1;green,1},looseness=1.0] (seqA5.south) to (g6);
\draw[line width=0.5,color={rgb:blue,1;green,1},looseness=1.0] (seqB5.south) to (g6);

\draw[line width=0.5,color=blue!100,looseness=1.0] (seqA6.south) to (g7);
\draw[line width=0.5,color=blue!100,looseness=1.0] (seqB6.south) to (g7);

\draw[line width=0.5,color={rgb:blue,1;purple,1},looseness=1.0] (seqA7.south) to (g8);
\draw[line width=0.5,color={rgb:blue,1;purple,1},looseness=1.0] (seqC5.south) to (g8);

\draw[line width=0.5,color=purple!100,looseness=1.0] (seqB7.south) to (g9);

\draw[line width=0.5,color=gray,looseness=1.0] (seqB2.south) to (g3);

\begin{comment}
\begin{scope}[node distance=0.5cm]
  \node[keynode,fill=red!100] (c1) [right = 1cm of seqA7] {1};
  \node[keynode,fill={rgb:red,1;orange,1}] (c2) [below of=c1] {2};
  \node[keynode,fill=orange!100] (c3) [below of=c2] {3};
  \node[keynode,fill=green!100] (c4) [below of=c3] {4};
  \node[keynode,fill={rgb:blue,1;green,1},text=white] (c5) [below of=c4] {5};
  \node[keynode,fill=blue!100,text=white] (c6) [below of=c5] {6};
  \node[keynode,fill={rgb:blue,1;purple,1},text=white] (c7) [below of=c6] {7};
  \node[keynode,fill=purple!100] (c8) [below of=c7] {8};
  \node[keynode,fill=gray] (c9) [below of=c8] {9};
\end{scope}
\end{comment}

\end{pgfonlayer}

%\draw[->] (uppercircle.south) -- (seq1.north);
%\draw[->] (seq1.east) -- (rightsquare.west);
%\draw[->] (rightsquare.south) .. controls +(down:7mm) and +(right:7mm) .. (lowercircle.east);
%\draw[->] (rightsquare.south) -- (lowercircle.east);
\end{tikzpicture}
\end{document}