### 2022-01-08 manuscript/fig_from_alignment_to_variation_graph.tex
get the first figure in

\vspace{-1em}
\begin{figure}[ht!]
%  \centering
   \includegraphics[width=\linewidth]{fig_induction.pdf}
   \caption{
   A visual description of variation graph induction.
   At top, the alignment graph model.
   %   At bottom, a sequence graph obtained by contracting

   %   At top, three sequences (text1, text2, text3) 
    }
    %{\includegraphics[width=\linewidth]{fig/OUP_First_SBk_Bot_8401}} %\includegraphics[width=5cm]{name}}
    %\caption{Performance evaluation of \textit{odgi build} when translating a 90-haplotype graph of human chromosome 6 into ODGI's native format.}
\end{figure}
\vspace{-1em}